---
title: "Highcharts : R User Day"
author: "Sam Hunter - NRS"
format: revealjs
editor: visual
---


```{r setup, include=FALSE}

# this handy little expression sources each script in the "scripts" folder
# if you're looking for code to make the highcharts check there.
for (script in list.files("scripts", full.names = TRUE)) {
  source(script, local = T)
}

```

## This presentation on Highcharts

A bit of background information to raise awareness and share our learning from using Highcharts in publications.

-   What is it?

-   Why use it?

-   Where has it been used?

-   What have we learned?

-   What are we working on?

## Reproducibility

### Source code for producing these slides is available on [github](https://github.com/DataScienceScotland/R-user-day-talks/tree/main/2023-12-06).

### Contact me: [sam.hunter@nrscotland.gov.uk](mailto:sam.hunter@nrscotland.gov.uk)

# Highcharts

## Highcharts

What is it?

-   Highcharts is a software library for data visualisation, written in pure JavaScript.

-   It is free for personal/non-commercial uses and paid for commercial applications.

-   The Scottish Government and NRS have licenses for publications.

## Highcharts

Why use it to make charts with?

-   Meant for HTML. To be viewed in a browser.

-   Lightweight. Charts are rendered in the page by the browser as vector graphics.

-   Interactive/ reactive. Charts change based on user interaction.

-   [Accessibility](https://www.highcharts.com/blog/posts/a11y/). Works with screen readers, charts can be sonified for visually impaired.

-   Customisation. Fine control over features; Colours, Fonts, Patterns, Download images/ data.

## R to HTML and JavaScript

But this is the R User Day... Why are you telling me about JS?

-   Works great embedded in HTML from R markdown and R Shiny.

-   The R package `highcharter` enables building charts entirely in R.

-   We don't need to know any JS to go from data to visualisation and still get all the benefits.

## Publications

Well if it's so good, where have highcharts been used in publications?

-   [Census outputs](https://www.scotlandscensus.gov.uk/2022-results/scotland-s-census-2022-rounded-population-estimates/)

-   [Registrar General's Annual Review](https://www.nrscotland.gov.uk/files/statistics/rgar/2022/scotlands-population-2022.html)

-   [Poverty Statistics](https://data.gov.scot/poverty/)

-   [Time Use in Scotland](https://data.gov.scot/time-use-scotland/time-use-2023/sex.html#introduction)


# What have we learned about producing highcharts?

## Straightforward to build and edit

-   Similar syntax to ggplot.

-   Good [documentation](https://www.highcharts.com/docs/index), [resources](https://jkunst.com/highcharter/) and [demos](https://dantonnoriega.github.io/ultinomics.org/posts/2017-04-05-highcharter-explainer.html).

-   A little JavaScript goes a long way. 

-   Great with RAPs for reports and publications.

# What Highcharts features have we used?

## [Line charts with tooltips](https://www.scotlandscensus.gov.uk/2022-results/scotland-s-census-2022-rounded-population-estimates/#section4)

```{r, echo=FALSE}
# 
plots[["POPULATION_ageing"]]

```


## [Bar charts with annotation](https://www.scotlandscensus.gov.uk/2022-results/scotland-s-census-2022-rounded-population-estimates/#section2)

```{r, echo=FALSE}
# 
plots[["census_records"]]

```

## [Choropleth Maps](https://www.scotlandscensus.gov.uk/2022-results/scotland-s-census-2022-rounded-population-estimates/#section5)

```{r, echo=FALSE}
# 
plots[["POPULATION_map"]]

```

Requires some geography files in [*geojson* format](https://github.com/martinjc/UK-GeoJSON/tree/master)

## [Two Chart Types with Flipped Axis](https://www.scotlandscensus.gov.uk/2022-results/scotland-s-census-2022-rounded-population-estimates/#section4)

```{r, echo=FALSE}
# 
plots[["pop_pyr_1921"]]

```

Horizontal axis is positive in both directions.

# What's next?

What other Highcharts features are we learning to use?

## Sonification

Sonification doesn't work inside this slideshow but does in standalone HTML knitted from R markdown.

[SONIFICATION DEMO]("../2023-12-06/Highcharts/sonification_demo.html")

# Pros and Cons

## Pros

-   Powerful for web based content.

-   Interactive by default.

-   Easy to get started.

-   Accessibility features.

-   Charts resize well.

-   Looks professional.

## Cons 

-   Paid license for publishing.

-   JavaScript.

-   Endless options.

-   Adding sonification isn't straightforward.

-   Hard to master.

# Alternatives

## ggplot + wrappers

### ggplotly / ggigraph

### Pros

builds off ggplot, easy to add on to existing code. FREE!

### Cons

Not so flexible, ggigraph isn't popular, small feature set for interactivity.

# How do I get started?

## Resources

Start with the...

Good [documentation](https://www.highcharts.com/docs/index), [resources](https://jkunst.com/highcharter/) and [demos](https://dantonnoriega.github.io/ultinomics.org/posts/2017-04-05-highcharter-explainer.html).

Further reading in the R Users' [sharepoint docs](https://scotsconnect.sharepoint.com/:w:/r/sites/rusers-community/_layouts/15/Doc2.aspx?action=edit&sourcedoc=%7BCDF2D182-A665-4C54-BD57-E5D4F045AD29%7D&wdOrigin=TEAMS-ELECTRON.null_ns.bim&wdExp=TEAMS-CONTROL&wdhostclicktime=1700743026986&web=1)

## Join the Community

-   Join the [RAP Yammer group](https://web.yammer.com/main/groups/eyJfdHlwZSI6Ikdyb3VwIiwiaWQiOiIxOTI2NDMyMzU4NSJ9)

-   Request to join the Teams "Highcharter Users Group - Org - NRS" [here](mailto:sam.hunter@nrscotland.gov.uk).

## Reproducibility

### Code for producing these slides is available on [github](https://github.com/DataScienceScotland/R-user-day-talks/tree/main/2023-12-06).

### Contact me: [sam.hunter@nrscotland.gov.uk](mailto:sam.hunter@nrscotland.gov.uk)

# Any Questions?
